Optimal Task Phasing for End-To-End Latency in Harmonic Systems and Semi-Harmonic Automotive Systems

### 2024-09-26 paper.tex
max harmonic

\begin{document}
	%% TITLE:
	\title{Optimal Task Phasing for End-To-End Latency in Harmonic and Semi-Harmonic Automotive Systems}

@@

@@

	Task systems are \emph{harmonic} if all task periods are integer multiples of one another.
	That is, for any two tasks $\tau, \tau' \in \Tbb$, we have $\frac{T_{\tau}}{T_{\tau'}} \in \Nbb$ or $\frac{T_{\tau'}}{T_{\tau}} \in \Nbb$.
	Harmonic task systems have the benefit that their release pattern is highly repetitive. As a result, the data propagation through these tasks is easier to analyze and to control.
	For the results of this work, weaker forms of harmonic are already sufficient.
	Specifically, in Section~\ref{sec:harmonic}, we consider \emph{max-harmonic} tasks, where 
	\begin{equation}
		\frac{T_{\mathit{max}}}{T_{\tau}} \in \Nbb
	\end{equation}
	for all $\tau \in \Tbb$, with $T_{\mathit{max}} := \max_{\tau\in \Tbb} T_{\tau}$. 
	For the results in Section~\ref{sec:automotive}, we consider semi-harmonic task sets with 
	\begin{equation}
		T_\tau \in \setof{1,2,5,10,20,50,100,200,1000}
	\end{equation}
	for all $\tau \in \Tbb$.
	These semi-harmonic task sets are common in the automotive domain~\cite{kramer2015real}.

@@

\section{Problem Definition}
\label{sec:problem_def}

	What is the best phasing? Nobody has looked into that yet.

\section{Optimal Task Phasing for Harmonic Systems and Max-Harmonic Systems}
\label{sec:harmonic}

	Indeed, the results of this section work even for a larger class of task systems.

	\begin{definition}[Max-Harmonic]
		A task system is called \emph{max-harmonic} if all periods divide the largest period, i.e., 
		\begin{equation}
			\frac{T_{\mathit{max}}}{T_{\tau_i}} \in \Nbb
		\end{equation}
		for all $\tau_i \in \Tbb$, where $T_{\mathit{max}} := \max_{\tau\in \Tbb} T_{\tau}$.
	\end{definition}

	\begin{lemma}[Optimality of task phasing]
		The end-to-end latency of $E$ under \emph{any} task phasing is lower bounded by $\sum_{i=1}^{n} T_{\tau_i} + \max_{i=1, \dots, n} T_{\tau_i}$.
		Hence, the task phasing from Theorem~\ref{thm:key} is optimal.
	\end{lemma}

@@

@@

>>>>

\begin{document}
	%% TITLE:
	\title{Optimal Task Phasing for End-To-End Latency in Harmonic Systems and Semi-Harmonic Automotive Systems}

@@

	\mario{TODO:
	- We use task set and task system interchangably, and say the task set/task system/system is harmonic. We need to make the language consistent.
	}

@@

	Task systems are \emph{harmonic} if all task periods are integer multiples of one another.
	That is, for any two tasks $\tau, \tau' \in \Tbb$, we have 
	\begin{equation}
		\frac{T_{\tau}}{T_{\tau'}} \in \Nbb \qquad \text{or} \qquad \frac{T_{\tau'}}{T_{\tau}} \in \Nbb.
	\end{equation}
	Harmonic task systems have the benefit that their release pattern is highly repetitive. As a result, the data propagation through these tasks is easier to analyze and to control.
	For the results of this work, weaker forms of harmonic are already sufficient.
	Specifically, in Section~\ref{sec:harmonic}, we consider \emph{max-harmonic} subsets $\Tbb' \subseteq \Tbb$, where 
	\begin{equation}
		\frac{\max_{\tau\in \Tbb'} T_{\tau}}{T_{\tau}} \in \Nbb
	\end{equation}
	for all $\tau \in \Tbb'$. 
	For the results in Section~\ref{sec:automotive}, we consider semi-harmonic task sets $\Tbb$ with 
	\begin{equation}
		T_\tau \in \setof{1,2,5,10,20,50,100,200,1000}
	\end{equation}
	for all $\tau \in \Tbb$.
	These semi-harmonic task sets are common in the automotive domain~\cite{kramer2015real}.

@@

% \section{Problem Definition}
% \label{sec:problem_def}

% 	What is the best phasing? Nobody has looked into that yet.

\section{Optimal Task Phasing for Harmonic Systems}
\label{sec:harmonic}

	As the motivating example demonstrates, there is a high potential of reducing the end-to-end latency by modifying the task phases. 
	In this section, we
	consider a cause-effect chain 
	\begin{equation}
		E = (\tau_1 \to\dots\to \tau_n),
	\end{equation}
	and
	explore optimal task phasing 
	that minimizes $\lat(E)$ if the system is harmonic.
	Actually, the results of this section work even for a larger class of task systems, in which the tasks 
	\begin{equation}
		\Tbb_E := \setof{\tau_1, \dots, \tau_n} \subseteq \Tbb
	\end{equation}
	of the cause-effect chain $E$ are \emph{max-harmonic}.

	\begin{definition}[Max-Harmonic]
		A task set $\Tbb' \subseteq \Tbb$ is called \emph{max-harmonic} if all periods divide the largest period, i.e., 
		\begin{equation}
			\frac{\max_{\tau\in \Tbb'} T_{\tau}}{T_{\tau}} \in \Nbb
		\end{equation}
		for all $\tau \in \Tbb'$.
	\end{definition}

	Naturally, if $\Tbb$ is harmonic, then $\Tbb$ and all its subsets $\Tbb' \subseteq \Tbb$ are also max-harmonic. 
	In particular, $\Tbb_E$ is max-harmonic.

	To develop an optimal task phasing strategy, we first formulate a general lower bound on the end-to-end latency.

	\begin{lemma}
		The end-to-end latency of $E$ under \emph{any} task phasing is lower bounded by $\sum_{i=1}^{n} T_{\tau_i} + \max_{i=1, \dots, n} T_{\tau_i}$.
	\end{lemma}

@@

	In the following, we develop a task phasing that ensures that the end-to-end latency reaches this lower bound for max-harmonic tasks.

	\mario{HERE}

@@

	\mario{Note: The lower bound is formulated for the general case. For max-harmonic tasks, we have shown that the lower bound is tight.}
